untitled

### 2016-04-30 appendix.tex
new appendices

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\\
  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2},\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \delta\bm{F}\cdot\hat{\bm{r}},\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}},
\end{align} 
with
\begin{align}
  \delta \bm{F} \equiv \frac{\dd\rho}{\rho^2}\nabla P -
  \frac{1}{\rho}\nabla\dd P.
\end{align}
Note that for the axisymmetric problem, $\dd\bm{F}$ is purely
meridional. 
From the linearized meridional momentum equations, we find 
\begin{align}
  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV = \int\left( \rho
  \kappa^2 |\dd v_r|^2 + \rho r\frac{\p \Omega^2}{\p z} \dd v_z \dd
  v_r^*  + \ii \rho \dd \bm{F}\cdot\dd\bm{v}^*
  \right)dV,
\end{align}
where the integral is taken over the volume of the fluid. Integrating
by parts and ignoring surface integrals, the last term is
\begin{align}
  \int \ii\sigma\rho \dd \bm{F}\cdot\dd\bm{v}^*dV &=\int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot \nabla P - \ii\sigma \dd
  \bm{v}^*\cdot\nabla \dd P  \right)dV\notag\\
&= \int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot\nabla P + \ii\sigma \dd
 P  \nabla\cdot \dd \bm{v}^*  \right)dV\notag\\
 & = \int\left[
   \left(
   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
   \bm{v} \cdot \nabla \ln{c_s^2}
   \right)\dd\bm{v}^*\cdot\nabla P  + \ii\sigma \dd
   P  \nabla\cdot \dd \bm{v}^*
   \right]dV\notag\\
 & = \int\left[
   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s\right) 
   \right]dV\notag\\
 & = \int\left\{
   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
   \dd\bm{v}\cdot\nabla\ln{c_s^2}\right]\nabla\cdot\left(P\dd\bm{v}^*\right)
   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s \right)
   \right\}dV \notag\\
   &=\int\left[
     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
     \right) -
     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right) 
     \right]dV,
\end{align}
where we recall $s\equiv\ln{(P/\rho)}$. Hence,
\begin{align}\label{integral_ex}
  &\sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
&=  \int\left\{
  \rho |\dd v_r|^2 \left(\kappa^2 - \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p r}\right)
  + \rho |\dd v_z|^2\left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p z}\right)
   + \rho \dd v_z \dd v_r^*\left(r\frac{\p\Omega^2}{\p z} -
  \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p z}\right) 
  + \rho \dd v_z^*\dd v_r \left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p r}\right)\right. \notag\\
&\phantom{==\int}\left.+
  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
  \right\}dV -\int P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.
\end{align}
Note that the coefficients of the mixed terms are equal owing to the
equilibrium state.

@@

>>>>

\begin{align}
  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{\rho},\\
  \ii\sigma\frac{\delta P}{P} &= \nabla\cdot\dd\bm{v} +
  \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{c_s^2},\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \delta\bm{F}\cdot\hat{\bm{r}},\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \delta\bm{F}\cdot\hat{\bm{z}},
\end{align} 
with
\begin{align}
  \delta \bm{F} \equiv \frac{\dd\rho}{\rho^2}\nabla P -
  \frac{1}{\rho}\nabla\dd P.
\end{align}
Note that for the axisymmetric problem, $\dd\bm{F}$ is purely
meridional. 
From the linearized meridional momentum equations, we find 
\begin{align}
  \sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV = \int\left( \rho
  \kappa^2 |\dd v_r|^2 + \rho r\frac{\p \Omega^2}{\p z} \dd v_z \dd
  v_r^*  + \ii\sigma \rho \dd \bm{F}\cdot\dd\bm{v}^*
  \right)dV,
\end{align}
where the integral is taken over the volume of the fluid. Integrating
by parts and ignoring surface integrals, the last term is
\begin{align}
  \int \ii\sigma\rho \dd \bm{F}\cdot\dd\bm{v}^*dV &=\int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot \nabla P - \ii\sigma \dd
  \bm{v}^*\cdot\nabla \dd P  \right)dV\notag\\
&= \int\left( \ii\sigma \frac{\dd
    \rho}{\rho}\dd\bm{v}^*\cdot\nabla P + \ii\sigma \dd
 P  \nabla\cdot \dd \bm{v}^*  \right)dV\notag\\
 & = \int\left[
   \left(
   \ii\sigma \frac{\dd P}{P} - \dd\bm{v}\cdot\nabla s + \dd
   \bm{v} \cdot \nabla \ln{c_s^2}
   \right)\dd\bm{v}^*\cdot\nabla P  + \ii\sigma \dd
   P  \nabla\cdot \dd \bm{v}^*
   \right]dV\notag\\
 & = \int\left[
   \ii\sigma\frac{\dd P}{P}\nabla\cdot\left(P\dd\bm{v}^*\right) +
   \left(\dd\bm{v}^*\cdot\nabla P\right) \left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s\right) 
   \right]dV\notag\\
 & = \int\left\{
   \left[\frac{1}{P}\nabla\cdot\left(P\dd\bm{v}\right) -
   \dd\bm{v}\cdot\nabla\ln{c_s^2}\right]\nabla\cdot\left(P\dd\bm{v}^*\right)
   +\left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v} \cdot \nabla
   \ln{c_s^2}   -  \dd\bm{v}\cdot\nabla s \right)
   \right\}dV \notag\\
   &=\int\left[
     \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2 -
     \left(\dd\bm{v}^*\cdot\nabla P\right)\left(\dd\bm{v}\cdot\nabla s
     \right) -
     P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right) 
     \right]dV,
\end{align}
where we recall $s\equiv\ln{(P/\rho)}$. Hence,
\begin{align}\label{integral_ex}
  &\sigma^2\int\rho\left(|\dd v_r|^2 + |\dd v_z|^2\right)dV \notag\\
&=  \int\left\{
  \rho |\dd v_r|^2 \left(\kappa^2 - \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p r}\right)
  + \rho |\dd v_z|^2\left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p z}\right)
   + \rho \dd v_z \dd v_r^*\left(r\frac{\p\Omega^2}{\p z} -
  \frac{1}{\rho}\frac{\p P}{\p
    r}\frac{\p s}{\p z}\right) 
  + \rho \dd v_z^*\dd v_r \left(-\frac{1}{\rho}\frac{\p P}{\p
    z}\frac{\p s}{\p r}\right)\right. \notag\\
&\phantom{==\int}\left.+
  \frac{1}{P}\left|\nabla\cdot\left(P\dd\bm{v}\right)\right|^2
  \right\}dV -\int P\left(\nabla\cdot\dd\bm{v}^*\right)\left(\dd\bm{v}\cdot\nabla\ln{c_s^2}\right)dV.
\end{align}
Note that the coefficients of the mixed terms are equal owing to the
equilibrium state.

@@

\section{Explicit expressions for pressure forces and its divergence}
Define
\begin{align}
  \bm{F} \equiv - \frac{\nabla P}{\rho}. 
\end{align}
For a locally isothermal equation of state, Eq. \ref{eos}, this
becomes
\begin{align}
  \bm{F} = c_s^2\nabla \epsilon - c_s^2(1-\epsilon)\nabla\ln{\rho} -
  (1-\epsilon)\nabla c_s^2,
\end{align} 
and its linearized form is
\begin{align}
  \dd\bm{F} = c_s^2\nabla\dd\epsilon + c_s^2\dd\epsilon\nabla\ln{\rho}
  - c_s^2(1-\epsilon)\nabla\left(\frac{\dd\rho}{\rho}\right) +
  \dd\epsilon\nabla c_s^2. 
\end{align}
The divergence of the above expressions are
\begin{align}
  \nabla\cdot\bm{F} &= 2 \nabla c_s^2\cdot\nabla\epsilon +
  c_s^2\nabla^2\epsilon  - \left[(1-\epsilon)\nabla c_s^2 -
    c_s^2\nabla\epsilon\right]\cdot\nabla\ln{\rho}
 - c_s^2(1-\epsilon)\nabla^2\ln{\rho} - (1-\epsilon)\nabla^2 c_s^2,\\
 \nabla\cdot\dd\bm{F} &= 2\nabla c_s^2\cdot\nabla\dd\epsilon +
 c_s^2\nabla^2\dd\epsilon + \left(\dd\epsilon \nabla c_s^2 +
 c_s^2\nabla\dd\epsilon\right)\cdot\nabla\ln{\rho} + c_s^2 \dd\epsilon
 \nabla^2\ln{\rho}   + \dd\epsilon\nabla^2 c_s^2 \notag\\
&\phantom{=}-c_s^2(1-\epsilon)\nabla^2\left(\frac{\dd\rho}{\rho}\right) - \left[(1-\epsilon)\nabla c_s^2 -
    c_s^2\nabla\epsilon\right]\cdot\nabla \left(\frac{\dd\rho}{\rho}\right).
\end{align}
For a power-law prescription for the radial temperature
profile $c_s^2\propto r^q$, we have
\begin{align}
  \nabla c_s^2 &= \frac{qc_s^2}{r}\hat{\bm{r}},\\
  \nabla^2 c_s^2 & = \left(\frac{q}{r}\right)^2 c_s^2. 
\end{align}

\section{Linearized dust evolution equation}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop \propto \frac{1}{\rho c_s}
\end{align}
\citep{price15}. 
Then the evolutionary equation for the dust-fraction $\epsilon$ is 
\begin{align}\label{dust_epstein}
  \frac{D\epsilon}{Dt} = \tstop c_s \nabla \cdot
  \left(\frac{\epsilon}{c_s}\bm{F}\right). 
\end{align} 
Linearizing Eq. \ref{dust_epstein},
\begin{align}
  s \dd\epsilon + \dd\bm{v}\cdot\nabla\epsilon &= \tstop
  \left(\dd\epsilon\nabla\cdot\bm{F} + \epsilon\nabla\cdot\dd\bm{F} +  
  \nabla\dd\epsilon\cdot\bm{F} + \nabla \epsilon\cdot\dd\bm{F} -
  \frac{1}{2}\dd\epsilon\bm{F}\cdot\nabla\ln{c_s^2} -
  \frac{1}{2}\epsilon\dd\bm{F}\cdot\nabla\ln{c_s^2} \right)
  \notag\\
  &\phantom{=} - \tstop \left(\epsilon\nabla\cdot\bm{F} +
  \nabla\epsilon\cdot\bm{F}-\frac{1}{2}\epsilon\bm{F}\cdot\nabla\ln{c_s^2}\right)\frac{\dd\rho}{\rho}.
\end{align}

\section{Conversion formulae}
The gas density $\rhog$ and dust-to-gas ratio $\tepsilon$ is related
to the total density $\rho$ and dust-fraction $\epsilon$ by
\begin{align}
  \ln{\rho} &= \ln{\rhog} + \ln{\left(1 + \tepsilon\right)},\\
  \nabla\ln{\rho} &= \nabla\ln{\rhog} + \frac{\nabla\tepsilon}{1+
    \tepsilon},\\
  \nabla^2\ln{\rho} &= \nabla^2\ln{\rhog} +
  \frac{\nabla^2\tepsilon}{1+\tepsilon} -
  \frac{\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^2}, 
\end{align}
and
\begin{align}
  \epsilon &= \frac{\tepsilon}{1+ \tepsilon},\\
 \nabla\epsilon & =
 \frac{\nabla\tepsilon}{\left(1+\tepsilon\right)^2},\\
 \nabla^2\epsilon &=
 \frac{\nabla^2\tepsilon}{\left(1+\tepsilon\right)^2} -
 \frac{2\left|\nabla\tepsilon\right|^2}{\left(1+\tepsilon\right)^3}. 
\end{align}
